Towards Pandemic-Scale Ancestral Recombination Graphs of SARS-CoV-2

### 2023-06-01 paper.tex
Updates from Overleaf

% Paragraph 2, what do we plot, specifically?
In this section we explore the detailed ancestral relationships between
recombinant parents by investigate the patterns of divergence between
them.
We focus on the Long ARG because it covers time
periods where substantial recombination is known to have occurred,
and contains samples from 33 Pango X lineages (i.e., those inferred
to have recombinant ancestry; see Section~\ref{sec:pango_x_lineages} for
further analysis).
Figure~\ref{fig:recomb_mrcas} shows the estimated date of the most recent
common ancestor (MRCA) of the parent nodes for each recombination breakpoint,
plotted against the divergence between these parents (i.e., the
total branch length from the parents to their MRCA in the trees to
the immediate left and right of the breakpoint).
As in Section~\ref{sec:breakpoint_intervals}, in these plots we
exclude breakpoints associated with ``singleton'' recombination nodes
(those ancestral to only one sample).
Larger points distinguish those breakpoints which occur in nodes
ancestral to 5 or more samples, comprising 316 breakpoints from
291 recombination nodes. The criterion of 5 descendants matches the
minimum number required to designate a new Pango recombinant lineage
\citep{Rambaut2020-dw}.
Note that as each plotted point represents a breakpoint, a recombination
node with more than one breakpoint (e.g., with 3 or more parents,
comprising ${\sim}10 \%$ of the
recombination nodes in
Figure~\ref{fig:recomb_mrcas}) will be represented by several points.

@@

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/supp_recombination_node_mrcas.pdf}
\caption{\label{fig:recomb_mrcas_voc_breakdown}  Divergence between parent
lineages for recombination events within and among different VoC categories.
There are 78 Alpha+Alpha recombination breakpoints corresponding to 75 recombination nodes
(25 breakpoints / 24 nodes with $\geq5$ descendant samples).
148 breakpoints from 142 nodes are Delta+Delta recombinations (45 breakpoints / 45 nodes with  $\geq5$ descendants),
and 148 breakpoints from 142 nodes are Omicron+Omicron (71 breakpoints / 68 nodes with $\geq5$ descendants).
The equivalent figures for Alpha+Delta are 9 / 8 (3 / 2),
for Alpha+Omicron are 2 / 1 (0 / 0),
and for Delta+Omicron are 20 / 13 (6 / 3).
Note that these plots do not include recombination breakpoints involving lineages that were not classified into these VoC categories.
}
\end{figure}

>>>>

% Paragraph 2, what do we plot, specifically?
In this section we explore the detailed ancestral relationships between
recombinant parents by investigating the patterns of divergence between
them.
We focus on the Long ARG because it covers time
periods where substantial recombination is known to have occurred,
and contains samples from 33 Pango X lineages (i.e., those inferred
to have recombinant ancestry; see Section~\ref{sec:pango_x_lineages} for
further analysis).
Figure~\ref{fig:recomb_mrcas} shows the estimated date of the most recent
common ancestor (MRCA) of the parent nodes for each recombination breakpoint,
plotted against the divergence between these parents (i.e., the
total branch length from the parents to their MRCA in the trees to
the immediate left and right of the breakpoint).
As in Section~\ref{sec:breakpoint_intervals}, in these plots we
exclude breakpoints associated with ``singleton'' recombination nodes
(those ancestral to only one sample).
Larger points distinguish those breakpoints which occur in nodes
ancestral to 5 or more samples, comprising 316 breakpoints from
291 recombination nodes. The criterion of 5 descendants matches the
minimum number required to designate a new Pango recombinant lineage
\citep{Rambaut2020-dw}.
Note that as each plotted point represents a breakpoint, a recombination
node with more than one breakpoint (e.g., with 3 or more parents,
comprising ${\sim}10 \%$ of the
recombination nodes in
Figure~\ref{fig:recomb_mrcas}) will be represented by several points.

@@

\begin{figure} \centering
\includegraphics[width=\textwidth]{figures/supp_recombination_node_mrcas.pdf}
\caption{\label{fig:recomb_mrcas_voc_breakdown}  Divergence between parent
lineages for recombination events within and among different VoC categories.
There are 78 Alpha+Alpha recombination breakpoints corresponding to 75 recombination nodes
(25 breakpoints / 24 nodes with $\geq5$ descendant samples).
148 breakpoints from 142 nodes are Delta+Delta recombinations (45 breakpoints / 45 nodes with  $\geq5$ descendants),
and 148 breakpoints from 142 nodes are Omicron+Omicron (71 breakpoints / 68 nodes with $\geq5$ descendants).
The equivalent figures for Alpha+Delta are 9 / 8 (3 / 2),
for Alpha+Omicron are 2 / 1 (0 / 0),
and for Delta+Omicron are 20 / 13 (6 / 3).
Note only recombination breakpoints involving lineages
classified into Alpha, Delta, and Omicron VoC categories are plotted above:
all other breakpoints are omitted.
}
\end{figure}
